Three Parameterizations of Rasch model
======================================

Here we demonstrate three alternative ways to represent the simple Rasch model for dichotomous
data. 

## Approach 1

Use fixed slope parameters set to 1, freely estimate all intercepts, and freely estimate the 
variance of the latent trait.

```{r}
library(mirt)
dat <- expand.table(LSAT6)
mod <- mirt(dat, 1, itemtype = 'Rasch')
print(mod)
coef(mod)
```

## Approach 2

Use freely estimate slope parameters but constrain them to be equal across all items, freely 
estimate all intercepts.

```{r}
model <- mirt.model('Theta = 1-5
                    CONSTRAIN = (1-5, a1)')
mod2 <- mirt(dat, model)
print(mod2)
coef(mod2)
```

## Approach 3

Fix slopes to 1, freely estimate intercepts subject to the constraint that $\sum_j d_j = 0$,
and freely estimate the latent mean and variance.

```{r}
model2 <- mirt.model('Theta = 1-5
                     MEAN = Theta
                     COV = Theta*Theta')

#view how vector of parameters is organized internally
sv <- mirt(dat, model2, itemtype = 'Rasch', pars = 'values')
sv[sv$est, ]

#constraint: create function for solnp to compute constraint, and declare value in eqB
eqfun <- function(p, optim_args) sum(p[1:5]) #could use browser() here, if it helps
solnp_args <- list(eqfun=eqfun, eqB=0, LB = c(rep(-15, 6), 1e-4))

mod3 <- mirt(dat, model2, itemtype = 'Rasch', optimizer = 'solnp', solnp_args=solnp_args,
			 pars=sv)
print(mod3)
coef(mod3)
(ds <- sapply(coef(mod3)[1:5], function(x) x[,'d']))
sum(ds)
```
